### 2021-10-19 docs/pap/pap.Rmd
compendium set-up

---
title: "Pre-Analysis Plan: The People as 'Volk' or 'Bürger'? The Implications of Ethnic and Civic Conceptions of the People for the Measurement of Populist Attitudes"
date: "`r format(Sys.time(), '%d %B, %Y - %X (%Z)')`"
output: 
  pdf_document:
    toc: yes
    number_sections: true
editor_options: 
  chunk_output_type: console
bibliography: references.bib
urlcolor: blue
linkcolor: blue
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \floatsetup[table]{capposition=top}
   - \usepackage{booktabs}
   - \usepackage{amsmath}
   - \usepackage{longtable}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{threeparttable}
   - \usepackage{threeparttablex}
   - \usepackage[normalem]{ulem}
   - \usepackage{makecell}
   - \usepackage{xcolor}
   - \usepackage[utf8]{inputenc}
---
```{r setup, include=FALSE}
## include this at top of your RMarkdown file for pretty output
## make sure to have the printr package installed: install.packages('printr')
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
#library(printr)
library(tidyverse)
library(kableExtra)
options(kableExtra.latex.load_packages = FALSE)
```

\newpage

# Expectations

**Ethnic Conception hypothesis** (_H1_):  All things equal, agreement with the constitutive items of the populist attitude construct containing a reference to 'the people' is lower for an ethnic conception of the 'the people' (i.e., das Volk) than for a civic conception of 'the people' (i.e., Bürger).e.

**Exclusive National Identity hypothesis** (_H2a_):  All things equal, agreement with the constitutive items of the populist attitude construct containing a reference to 'the people' is higher for an ethnic conception of 'the people' (i.e., das Volk) than for a civic conception of 'the people' (i.e., Bürger) when the respondent adheres to an exclusive national identity.

**Far Right Party ID hypothesis** (_H2b_): All things equal, agreement with the constitutive items of the populist attitude construct containing a reference to 'the people' is higher for an ethnic conception of 'the people' (i.e., das Volk) than for a civic conception of 'the people' (i.e., Bürger) when the respondent votes for a far right party.

# Research Desing and Protocol

## Sample
We will conduct the survey experiment in Germany in October 2021 -- approximately three weeks after the general elections of 26 September 2021.
The sample, recruited through [Respondi](https://www.respondi.com/EN/), will consist of 8,000 participants (based on the power analysis presented in Figure \ref{fig:power}) of 18 years and older.
Respondi works with opt-in respondents, so we have implemented quota on age, gender, and education.
Moreover, we measure some more demographic background variables (see [Section 3.2](\#control-variables)).
Balance checks will be conducted to demonstrate whether certain categories are over represented in a certain experimental group.
The study has been approved by the [Research Ethics Review Committee](https://fsw.vu.nl/nl/onderzoek/research-ethics-review/index.aspx) of the \textit{Vrije Universiteit Amsterdam} (see the approval [here](LINK)).
To ensure good quality of our data, two attention checks (discussed in more detail in [Section 3.3](\#attention-checks)) are included.
Each respondent failing the attention check will be excluded and replaced with another 'good' response.

## Experimental Protocol
The study is conducted online and in German.
Participants are told that they are taking part in a survey to get an overview of how German people form their views on politics.
After reading an informed consent message participants are forwarded to the main questionnaire (or the survey will be terminated if they do not agree to the consent form).

First, participants complete a set of demographic variables (i.e. income, employment, and degree of urbanization).
This block ends with one of the two attention checks included in this survey.
When participants fail this attention check, a warning appears asking them to read the question again
carefully  and  to  answer  again.
Thereafter, a set of background variables are asked on their stances on the political issues, some are used in the experiment (i.e. top tax, speed limit, education, legalization of Cannabis), their political interest and knowledge, as well as on their ideological position -- the full codeboek can be viewed [here](#).[^1]
Before entering the experimental condition, respondents get a second attention check.
Only when they have answered it correctly, they enter the first round of the experiment -- if they fail to answer the question correctly, they are thrown out of the survey.
The stimuli in the experiment are to translate the word _people_ into it's civic conception (i.e. _Bürger_) or it's ethnic conception (i.e. _Volk_) for the items 2, 3, and 4 of the populism attitude scale [@akkerman2014populist].

[^1]: Participants are also part of another experiment -- see details [here](https://github.com/MarikenvdVelden/willingness-to-accept-compromises).

## Power Analysis
As detailed in [Section 4](#analysis), we conduct an OLS regression for each of the dependent variables (i.e. items 2--4): a) "The people, not politicians, should make the most important political decisions"; b) "The politicians in the German Bundestag need to follow the will of the people"; and c) "The political differences between elites and the people are greater than the differences among citizens".
  The manipulation, i.e. the civic conception (i.e. _Bürger_) or th ethnic conception (i.e. _Volk_) as well as respondents national identity and far right party ID are the main independent variables.

To calculate power for the hypotheses, the R package `DeclareDesign` is used [@blair2019declaring].
The effect sizes are between $b = 0.2$ and $b = 0.1$ -- i.e. a small effect visualized by the purple and blue lines in Figure \ref{fig:power1}.
The hypothesis are directional, Figure \ref{fig:power1} therefore displays one-tailed tests with $\alpha = 0.05$.
The power analysis shows that testing hypotheses 1 and 2 requires a sample size of $8,000$ participants (x-axis) to reach very high levels of power, i.e. $>95\%$, (indicated by the black dashed line in Figure \ref{fig:power}).

```{r power-example, eval = FALSE, echo = TRUE}
  source(here::here("docs/pre-analysis-plan/poweranalysis.R"))
```

```{r power, fig.align = 'center', out.width = "100%", fig.cap = "\\label{fig:power1}Power Analysis", fig.topcaption=TRUE}
  knitr::include_graphics(here::here("report/figures/power_analysis.png"))
```

# Measures

## Dependent Variables
We rely on the three items in the populism scale that mention the people: a) "The people, not politicians, should make the most important political decisions"; b) "The politicians in the German Bundestag need to follow the will of the people"; and c) "The political differences between elites and the people are greater than the differences among citizens". All items are measured on a 5-point Likert-scale: `very much disagree`, `disagree`, `neutral`, `agree`, `very much agree`.

See Table \ref{tab:dv} for the exact phrasing of the question.

```{r dv, topcaption=TRUE}
d <- tibble(Variable = c("Item 2 (DV1)",
                         "Item 3 (DV2)",
                         "Item 4 (DV3)"), 
            `Wording ENG` = c("The people, not politicians, should make the most important political decisions","he politicians in the German Bundestag need to follow the will of the people", "The political differences between elites and the people are greater than the differences among citizens"),
            `Wording DE` = c("[Die Bürger/Das Volk], und nicht die Politiker, sollte(n) die wichtigsten politischen Entscheidungen treffen", "Die Abgeordneten des Deutschen Bundestags müssen dem Willen [der Bürger/des Volkes] Folge leisten", "Die politischen Unterschiede zwischen den Eliten und [den Bürgern/dem Volk] sind größer als die Unterschiede (zwischen Bürgern)"))

kbl(d, booktabs =T, caption = "\\label{tab:dv}Survey Questions - DV") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "7cm") %>%
  column_spec(3, width = "7cm")

```

## Treatment
Respondents are randomly assigned to the populism items translated into German using either the civic conception (i.e. _Bürger_) or the ethnic conception (i.e. _Volk_).

&nbsp;

## Control Variables

As control variables, the following _demographics_ are measured: gender, age, education, geographical region, level of urbanness, vote choice in the 2021 parliamentary elections, employment, and income.
For the analysis, only variables that are unbalanced over the experimental conditions will be included.
Table \ref{tab:demographics} gives an overview of the questions asked in the survey as well as their English translations.

- _Gender_ is measured as `sex` based on the advice of Kieskompas. The answer categories are `Male` (value of `1`), `Female` (value of `0`), and `No answer` (value of `999`).

- _Age_ is measured using 6 categories: `17 or younger`, `18--29`, `30--39`, `40--49`, `50--59`, `60--74`.

- _Education_ is measured as the highest successfully completed level of education, recoded into four categories: `low`, `middle`, `high`, and `none`.
I create dummy variables for each level of education with the lowest category as base category.

- _Urbanness_ Respondents are asked for what the type of location they live: `big city`, `suburb of a large city`, `middle-szed or small city`, `rural village`, `detached house in the countryside`, `don't know/won't say`.

- _Vote Recall_ Respondents were asked which party they voted for in the last election with their second vote. Answering categories were: `CDU/CSU`, `SPD`, `AfD`, `FDP`, `DIE LINKE`, `BÜNDNIS 90/DIE GRÜNEN`, `Other party:`, `I didn't vote/wasn't eligble to vote`, `Don't know/won't say`.

- _Employment_ Respondents were asked which category of employment -- `Full-time employed`, `Part-time employed`, `Entrepreneur`, `Unemployed and searching for a job`, `Unemployed and not searching for a job or incapacitated`, `Housewife/Househusband or else`, `Retired`, `Student or full-time education` -- applied most to them.

- _Income_ Respondents were questioned on their monthly income in bins of €500 -- `€500 or less`, `€501-€1000`, `€1001-€1500`, `€1501-€2000`, `€2001-€2500`, `€2501-€3000`, `€3001-€3500`, `€3501-€4000`, `€4501-€7500`, `€7501 or more` -- as well as giving them the options of `won't  say` and `don't know`.

- _Geographical region_ is measured using the same questions as the [German Longitudinal Election Study](https://www.gesis.org/wahlen/gles), asking respondents where they are born and currently live, differentiating between the different German provinces (_Bundesl{\"a}nder_): Baden-W{\"u}rtemberg, Bavaria, Berlin, Brandenburg, Bremen, Hamburg, Hessen, Lower Saxony, Mecklenburg-Western Pomerania, North Rhine-
Westphalia, Rhineland-Palatinate, Saarland, Saxony, Saxony-Anholt, Schleswig-Holstein, and Thuringia.

```{r demographics, topcaption=TRUE}
d <- tibble(Variable = c("gender (D1)", "age (D2)", "education (D3)",
                         "geographical-region (D4)", "vote-recall2 (D6)", "employment (D7)",
                         "income (D8)", "living-place (D9)", "birth-place (D10)"), 
            `Wording ENG` = c("Which gender do you feel you belong to?", "To which of the following age groups do you belong?",
                              "What is your highest educational qualification?", 
                             "Which of the following categories best describes where you live?",
                             "Which party did you vote for with your second vote in the last federal election in September 2021?",
                             "Now we would like to ask you something about your employment. Which of this list applies to you?",
                              "What is the total monthly net income of your household? This refers to the sum that remains after deducting taxes and social security contributions.",
                             "In which federal state or on the territory of which present federal state do you currently live?",
                              "And in which federal state or on the territory of which present federal state were you born?"),
            `Wording DE` = c("Welchem Geschlecht fühlen Sie sich zugehörig?", 
                             "Zu welcher der nachfolgenden Altersgruppen gehören Sie?",
                             "Was ist Ihr höchster Bildungsabschluss?",
                             "Welche der folgenden Kategorien beschreibt am besten, wo Sie wohnen?",
                             "Welche Partei haben Sie bei der letzten Bundestagswahl im September 2021 mit Ihrer Zweitstimme gewählt?",
                             "Nun möchten wir Sie gerne etwas zu Ihrer Erwerbstätigkeit fragen. Was von dieser Liste trifft auf Sie zu?",
                             "Wie hoch ist das monatliche Netto-Einkommen Ihres Haushaltes insgesamt? Gemeint ist die Summe, die nach Abzug von Steuern und Sozialversicherungsbeiträgen übrig bleibt.",
                             "Im welchen Bundesland bzw. auf dem Gebiet welches heutigen Bundeslandes wohnen Sie derzeit?",
                             "Und in welchem Bundesland bzw. auf dem Gebiet welches heutigen Bundeslandes wurden Sie geboren?"))

kbl(d, booktabs =T, caption = "\\label{tab:demographics}Survey Questions - Demographics") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "7cm") %>%
  column_spec(3, width = "7cm")

```

In addition, pre-treatment, respondents' ideological position, political knowledge, interest, and attitude towards national identity are measured (see Table \ref{tab:pret}).
Those variables will only be included in the analyses if balance checks indicate they are necessary.
Moreover, the variables will be used to explore heterogeneous relationships.

- _Ideological position_ is measured using an 11-point scale ranging from left (`0`) to right (`10`).

- _Exclusive national identity_ is measured with four statements ("being born in Germany", "having German ancestors", "being able to speak German", "adhering to German traditions and customs") on a 5-point Likert-scale: `very much disagree`, `disagree`, `neutral`, `agree`, `very much agree`. If `Cronbach's` $\alpha =>0.8$, we will construct an additive scale for this measure.
When `Cronbach's` $\alpha <0.8$, we will add the statement as seperate dependent variables in the multiverse analysis.

```{r pret, topcaption=TRUE}
d <- tibble(Variable = c("political-knowledge1 (PT4)", "political-knowledge2 (PT5)", "political-knowledge3 (PT6)", "political-interest (PT7)","RILE (PT8)", "Exclusive national identity (HT4)", "Exclusive national identity (HT4-1)", "Exclusive national identity (HT4-2)", "Exclusive national identity (HT4-3)", "Exclusive national identity (HT4-4)" ), 
            `Wording ENG` = c("In the federal election, you have two votes, a first vote and a second vote. How is that actually, which of the two votes is decisive for the distribution of seats in the Bundestag?",
                             "Now we would like to know from you, from what percentage of the second votes a party can send deputies in any case in the Bundestag?",
                             "And can you say approximately what the current unemployment rate is in Germany? Is it lower or higher than 10 percent?", 
                             "Once speaking in general terms: How interested are you in politics - very strongly, strongly, moderately, less strongly, or not at all?",
                            "In politics, people often talk about 'left' and 'right.' Where would you classify yourself?",
                            "Some people think that the following points are important to be truly German. Others do not think they are important. How important do you think the following points are to being German?", "being born in Germany", "having German ancestors", "being able to speak German", "adhering to German traditions and customs"),
            `Wording DE` = c("Bei der Bundestagswahl haben Sie ja zwei Stimmen, eine Erststimme und eine Zweitstimme. Wie ist das eigentlich, welche der beiden Stimmen ist ausschlaggebend für die Sitzverteilung im Bundestag?",
                             "Jetzt möchten wir gerne von Ihnen wissen, ab wie viel Prozent der Zweitstimmen eine Partei auf jeden Fall Abgeordnete in den Bundestag entsenden kann? ",
                             "Und können Sie ungefähr sagen, wie hoch die derzeitige Arbeitslosenquote in Deutschland ist? Ist sie niedriger oder höher als 10 Prozent?", 
                             "Einmal ganz allgemein gesprochen: Wie stark interessieren Sie sich für Politik – sehr stark, stark, mittelmäßig, weniger stark oder überhaupt nicht?",
                            "In der Politik reden die Leute häufig von 'links' und 'rechts'. Wo würden Sie sich selbst einordnen?",
                            "Manche Leute meinen, dass die folgenden Punkte wichtig sind, um wirklich deutsch zu sein. Andere halten diese nicht für wichtig. Für wie wichtig halten Sie die folgenden Punkte, um deutsch zu sein?", "in Deutschland geboren sein", "deutsche Vorfahren haben", "deutsch sprechen können", "sich an deutsche Traditionen und Gepflogenheiten halten"))

kbl(d, booktabs =T, caption = "\\label{tab:pret}Survey Questions - PreTreatment Questions") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "7cm") %>%
  column_spec(3, width = "7cm")
```

- _Political knowledge_ is measured with three items. First, respondents are asked which of the two votes is determinative for the seat distribution in parliament: `first vote`, `second vote`, `both`, `don't know`. Secondly, respondents are asked about the electoral threshold. Last, respondents are asked whether or not the unemployment rate is over ten procent.

- _Political interest_ is measured by asking people how strongly they are interested in politics using a 5-point Likert-scale (`very strong`, `strong`, `medium`, `less strong`, `not at all`) and a `don't know/won't say` option.

## Attention Checks
We include two attention checks in the survey.
The first one is after the demographic covariates, the second one is asked just before respondents enter the round of the experimental treatments.
The attention checks are taken from @berinsky2014separating and adapted to the German context by the authors.
If a respondents fails the first attention check, a warning appears and the respondent can only continue with the survey once the respondent has correctly answered the question correctly.
The second attention check also has a warning -- meaning that respondents have to select two options -- but if they fail to correctly pass the check, they are excluded.
Each excluded respondent due to failing an attention check is replaced with another ``good respondent".

**Attention Check 1**  When a big news story breaks people often go online to get up-to-the-minute details on what is going on.
We want to know which websites people trust to get this information.
We also want to know if people are paying attention to the question.
To show that you have read this much, please ignore the question and select BILD-Zeitung and Süddeutsche Zeitung as your two answers.
When there is a big news story, which is the one news website you would visit first? (Please only choose one).
Eight (German) news outlets are provided to choose from.
Respondents pass the attention check if they select _BILD-Zeitung_ and _Süddeutsche Zeitung_.

**Attention Check 2**:
We would like to get a sense of your general preferences.
Most  modern  theories  of  decision  making  recognize  that  decisions  do  not take  place  in  a  vacuum.
Individual  preferences  and  knowledge,  along  with  situational variables can greatly impact the decision process.
To demonstrate that you've read this much,  just  go  ahead  and  select  both  red  and  green  among  the  alternatives  below,  no matter what your favourite color is.  Yes, ignore the question below and select both of those options.
What is your favourite color?
Six colors are provided to choose from, respondents pass the attention check if they select red and green.

## Exclusion Criteria
Participants are required to respond to each question.
Participants who fail the second attention check will be excluded but replaced by another participant.

# Analysis
We test the hypotheses formulated in [Section 1](#expectations) by fitting linear multivariate regressions separately for the two issues.
In each model, we will estimate the coefficient for whether the party compromised or remained steadfast (H1), whether or not the coalition talks continued (H2), and the interaction of compromise and the outcome (H3).
We will only add control variables in the analyses that are unbalanced, as explained in [Section 4.1](#balance-checks).

## Balance Checks
We will conduct a balance test based on demographics (age, gender, education, geographical region, level of urbanness,employment, and income), vote choice in the 2021 parliamentary elections, ideological self-placement, political knowledge, political interest, party identity strength, and positions on and salience of the three issues, using the `cobalt` R package [@greifer2021]. 
If the groups are unbalanced on one of these variables -- i.e. standardized mean differences $<0.05$ -- I will add the covariates to the analyses.
I will use the code below to conduct the balance tests (see [here](https://github.com/MarikenvdVelden/compromise-punish/tree/main/src/analysis/balance-test.R) for the R script).

```{r balance, eval=FALSE, echo = TRUE}
covs <- d %>%
  mutate(treatment = populism) %>%
  select(populism, D1:D10, PT1:PT8, HT4)

balanced <-bal.tab(treatment ~ factor(D1) + D2 +
                     factor(D3) + D4 + D5 + 
                     factor(D6) + D7 + D8 + D9 +
                     D10 + PT4 + PT5 + PT6 + PT7 + 
                     PT8 + HT4,
                     data = covs,
          thresholds = c(m = 0.05))[[1]] 
```

## Hypothesis 1
We test the _Ethnic Conception hypothesis_, we will conduct a difference in means for the group who received the ethnic and the civic question wording, for all populist items containing a reference to _the people_ -- see code chunck below.

```{r h1, eval = FALSE, echo = TRUE}
source(here::here("src/lib/functions.R"))

df %>% 
  mutate(ethnic == if_else(populist=="Volk", "Ethnic", "Civic")) %>%
  select(`POST-2`, `POST-3`, `POST-4`, ethnic) %>%
  gather(key = variable, value = value, -ethnic) %>%
  group_by(ethnic, variable) %>%
  summarise(value = list(value)) %>% 
  spread(ethnic, value) %>% 
  group_by(variable) %>% 
  mutate(p_value = t.test(unlist(Ethnic), unlist(Civic))$p.value,
         t_value = t.test(unlist(Ethnic), unlist(Civic))$statistic)
```

## Hypothesis 2a

We test the _Exclusive National Identity hypothesis_ using the equation and the code chunk below.
The coefficient of $\beta_{1}$ denotes the difference in populist attitude for civic and ethnic conceptions of the people.
The coefficient of $\beta_{2}$ denotes the difference in populist attitude for different levels of national identity.
The coefficient of $\beta_{3}$ is the interaction of the  $\beta_{1}$ and  $\beta_{2}$.
If the Average Marginal Effect of $\beta_{1}$ is positive and statistically significant for higher values of $\beta_{2}$ (i.e. stronger national identity), then H2a will be confirmed.

$$\begin{aligned}
    \hat{Y}_{item} = \beta_{0} + \beta_{1}Ethnic  +
    \beta_{2}National \ Identity + \beta_{3} Ethnic * National \ Identity + \nonumber \\
    \beta_{4}Unbalanced \ Covariates + \varepsilon
  \end{aligned}$$

```{r h2a, eval = FALSE, echo = TRUE}
source(here::here("src/lib/functions.R"))

df <- d %>% 
  mutate(ethnic == if_else(populist=="Volk", "Ethnic", "Civic"))  %>%
  select(`POST-2`, ethnic, HT4)
h2a_1 <- lm(`POST-2` ~  ethnic * HT4,
                   data = df)

df <- d %>% 
  mutate(ethnic == if_else(populist=="Volk", "Ethnic", "Civic"))  %>%
  select(`POST-3`, ethnic, HT4)
h2a_2 <- lm(`POST-3` ~  ethnic * HT4,
                   data = df)

df <- d %>% 
  mutate(ethnic == if_else(populist=="Volk", "Ethnic", "Civic"))  %>%
  select(`POST-4`, ethnic, HT4)
h2a_3 <- lm(`POST-4` ~  ethnic * HT4,
                   data = df)
```

## Hypothesis 2b

We test the _Far Right Party ID hypothesis_ using the equation and the code chunk below.
The coefficient of $\beta_{1}$ denotes the difference in populist attitude for civic and ethnic conceptions of the people.
The coefficient of $\beta_{2}$ denotes the difference in populist attitude for whether or not a respondent has voted for a far right party.
The coefficient of $\beta_{3}$ is the interaction of the  $\beta_{1}$ and  $\beta_{2}$.
If the Average Marginal Effect of $\beta_{1}$ is positive and statistically significant for higher values of $\beta_{2}=1$ (i.e. voted for a radical right party), then H2b will be confirmed.

$$\begin{aligned}
    \hat{Y}_{item} = \beta_{0} + \beta_{1}Ethnic  +
    \beta_{2}Far \ Right \ Party \ ID + \beta_{3} Ethnic * Far \ Right \ Party \ ID + \nonumber\\
    \beta_{4}Unbalanced \ Covariates + \varepsilon
  \end{aligned}$$

```{r h2b, eval = FALSE, echo = TRUE}
source(here::here("src/lib/functions.R"))

df <- d %>% 
  mutate(afd_vote == if_else(`D6-3`==1, 1, 0),
         ethnic == if_else(populist=="Volk", "Ethnic", "Civic"))  %>%
  select(`POST-2`, ethnic, afd_vote)
h2a_1 <- lm(`POST-2` ~  ethnic * afd_vote,
                   data = df)

df <- d %>% 
  mutate(afd_vote == if_else(`D6-3`==1, 1, 0),
         ethnic == if_else(populist=="Volk", "Ethnic", "Civic"))  %>%
  select(`POST-3`, ethnic, afd_vote)
h2a_2 <- lm(`POST-3` ~  ethnic * afd_vote,
                   data = df)

df <- d %>% 
  mutate(afd_vote == if_else(`D6-3`==1, 1, 0),
         ethnic == if_else(populist=="Volk", "Ethnic", "Civic"))  %>%
  select(`POST-4`, ethnic, afd_vote)
h2a_3 <- lm(`POST-4` ~  ethnic * afd_vote,
                   data = df)
```

## Statistical Significance
All the hypotheses are directional, and therefore all of the tests will be one-tailed.
We will use an $\alpha$-value of $0.05$ as the value for statistical significance in all models above.

\newpage
# References