### 2022-09-09 report/accepted-manuscript.Rmd
final version

---
output: 
  stevetemplates::article:
    fig_caption: true
bibliography: references.bib
urlcolor: black
linkcolor: black
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \floatsetup[table]{capposition=top}
   - \usepackage{booktabs}
   - \usepackage{longtable}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage[para,online,flushleft]{threeparttable}
   - \usepackage{threeparttablex}
   - \usepackage[normalem]{ulem}
   - \usepackage{makecell}
   - \usepackage{xcolor}
   - \usepackage{hyperref}
   - \usepackage{array}   
   - \usepackage{caption}
   - \usepackage{graphicx}
   - \usepackage{epstopdf}
   - \usepackage{siunitx}
   - \usepackage{multirow}
   - \usepackage{hhline}
   - \usepackage{calc}
   - \usepackage{tabularx}
   - \usepackage{fontawesome}
   - \usepackage{amsthm}
   - \newtheorem{hypo}{Hypothesis}
biblio-style: apsr
title: "The People as 'Volk' or 'Bürger'? The Implications of Ethnic and Civic Conceptions of the People for the Measurement of Populist Attitudes"
author:
  - name: Maurits Meijers
  - affiliation: Radboud University Nijmegen
  - name: Mariken A.C.G. van der Velden
  - affiliation: Vrije Universiteit Amsterdam
thanks: "Replication files are available on the author's Github account (https://anonymous.4open.science/r/wording-experiment-populist-attitudes-F3B2/README.md). **Current version**: `r format(Sys.time(), '%B %d, %Y')`;"
anonymous: FALSE
abstract: "Populists believe in the sovereignty of the people. Yet, the people can be construed in ethnic terms or in civic terms. Using a novel wording experiment in Germany (N=7,034), we examine whether ethnic or civic conceptions of 'the people' affect respondents' adherence agreement to key populist attitude items. '_Volk_' in German has an ethno-nationalist connotation and has been used throughout history to signify 'ethnic Germans'. The concept '_Bürger_', or citizens, by contrast, lacks an ethnic undercurrent. Similar ethnic connotations of 'the people' are also common in other languages. We find that there are statistically significant differences between items framed in an ethnic and a civic manner -- and that this differs per item. This relationship is significantly moderated by respondents' degree of exclusive national identity and voting behaviour for the radical right. Our findings suggest that the way in which the people is conceptualized has important implications for the measurement of populist attitudes. When populist attitudes are measured with an ethnic understanding of the 'the people', the construct is biased toward right-wing populism, inhibiting the measurement of populism as a `thin-ideology'. Moreover, we demonstrate the importance of careful translations in comparative research, since some translated synonyms carry different meanings, and thereby change the concept under investigation."

keywords: "Wording Experiment, Populist Attitudes, Scaling, Measurement Invariance, Populism"
geometry: margin=1in
mainfont: cochineal
fontsize: 11pt
params:
  anonymous: ""
  doublespacing: ""
doublespacing: TRUE
endnote: no
pandocparas: TRUE
sansitup: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message=FALSE, warning=FALSE)
options(kableExtra.latex.load_packages = FALSE)
options(scipen = 1, digits = 2)

# Uncomment below if you want code captions
# oldSource <- knitr::knit_hooks$get("source")
# knitr::knit_hooks$set(source = function(x, options) {
#   x <- oldSource(x, options)
#   x <- ifelse(!is.null(options$code.cap), paste0(x, "\\captionof{chunk}{", options$code.cap,"}"), x)
#   ifelse(!is.null(options$ref), paste0(x, "\\label{", options$ref,"}"), x)
# })
# Add `chunkcaption: TRUE` to YAML as well.
```

# Introduction
Populism puts 'the people' central in political decision-making. 
Its central claim is that political power should reside with the people, as opposed to the elites [@canovan1999trust; @mudde2004].
To study populist beliefs at the individual level, scholars recently developed the populist attitude scale [@akkerman2014; @hawkins2012; @castanho2020]. 
Research on both the individual and the party level has found that populism is a diverse phenomenon, applicable to inclusive left-wing ideologies and exclusionary right-wing ideologies [@meijers2021; @vanhauwaert2018]. 
Indeed, populism is thought to be a thin-centred ideology that is usually attached to 'full' ideologies, such as socialism or nationalism [@mudde2004]. 
As a result, the concept of populism remains agnostic about _who_ constitutes the people.

In civic terms, we can think of the people as 'citizens' [e.g., @almond1963]. 
In ethnic terms, 'the people' can refer to a more cultural understanding of _folk_. 
While in some  cultural contexts, there is no strong semantic difference between 'the people' and 'citizens' (e.g. in the English-speaking world), in many other contexts, 'the people' (i.e., _das Volk_) has a strong ethnic connotation. For example, in Italian _il popolo_ has a stronger ethnic character than _i cittadini_ or _le persone_. In Polish the term _naród_ signifies the 'nation' as an ethnic group, compared to _społeczeństwo_ ('the public') [@zubrzycki2001we]. While difference is particularly outspoken in German, the difference exists in more Germanic, Slavic and Roman languages.
In the 1800s, the term _Volk_ and _völkisch_ became strongly associated with an ethnic and racial conception of 'Germanness'. 
After World War I, the terms _Volk_ and _Volksdeutsche_ were used to describe a ethnically and culturally homogeneous German diaspora, who were citizens of different states after the Versailles Treaty of 1919 [@gosewinkel2003]. _Das deutsche Volk_ was also used by the Nazis to delineate ethnic Germans from non-ethnic Germans. In 1935, poet and playwright Bertold Brecht therefore called for the use of _Bevölkerung_ ('population' or 'the public') instead of _Volk_ to to avoid the Nazis' "foul mysticism" [@brecht1993funf].

Scales of populist attitudes have mostly relied on ethnic conceptions of the people [@akkerman2014; @hawkins2012; @castanho2020; @vanhauwaert2020; @schulz2018]. 
At the same time, populist attitude scales have often been inconsistent in juxtaposing the elites with 'the people' or with 'citizens'. 
For instance, the scale proposed by @akkerman2014 primarily refers to 'the people'. 
Yet one item in the @akkerman2014 scale contrasts 'the citizen' with 'politicians' (`POP4`) - an item that nonetheless contributes significantly to the latent construct of populism [@vanhauwaert2020]. It is also likely that differences arise from inconsistencies across translations of they original survey items. For instance, 'the people' has been translated differently in different translations of the populist attitude scale. 
For instance, in the French context 'the people' has been translated as _le peuple_ in some items, but as _les citoyens_ in other items [@ivaldi2018]. 
Yet another study of French citizens refrains from using either _le peuple_ or _les citoyens_, choosing the more neutral _les gens_ instead [@castanho2020]. 
In Italian, 'the people' is sometimes referred to as _il popolo_, and sometimes as _le persone_ [@castanho2020]. In Polish, 'the people' has been measured as _społeczeństwo_ (`the public') and the more ethnic _naród_.

The assumption is that the different conceptions of the people (i.e. civic or ethnic) functions similar -- also known as measurement invariance or measurement equivalence.[^1]
In this paper, we test the degree to which the populist attitude scale is invariant across civic and ethnic concpetions of the people: 
That is, do people exposed to a civic conception of the people have a similar populist attitude score as people exposed to an ethnic one?
This is important because populist attitudes aim to operationalize populism as a `thin ideology' as propagated by @mudde2004. This perspective posits that populism is usually combined with other, left- or right-wing, ideologies.
Hence, the populist attitude scale should capture populist ideation across different ideological predispositions. In a similar vein, @castanho2020 criticize populist items scales which include nativist items.

Using a wording experiment in Germany (`N=7,034`), we examine measurement invariance across ethic and civic conception of `the people'.
First, we study whether an ethnic or a civic conception of the people affects respondents' agreement with the two items of the populist attitude scale that directly juxtapose 'the people' with the elites -- i.e. assessing structural, metric, or weak invariance [@hirschfeld2014improving].[^2] 
We further examine this structural invariance, by testing whether far right voters and respondents with a exclusive national identity respond differently to ethnic conceptions of the people. 
In addition, we examine whether the differences found affect the overall latent construct of populist attitudes, looking whether the intercept and the error terms are invariant (i.e. scalar and strict invariance) [@hirschfeld2014improving]. 
We show that the populist attitudes scale based on a civic conception is not the measurement equivalent of the one using the ethnic conception.
The assumption of measurement invariance is thus not met. 
Using an civic conception leads to higher scores on the populist attitude scale compared to using an ethnic conception.
This suggest that the way in which the people is conceptualized has important implications for the measurement of populist attitudes as well as for its explanatory power. Our findings also suggest that political scientists should pay much more attention to the translations they commission for their surveys. While new survey items are carefully constructed, translations are often not put to the same scrutiny. We demonstrate that using some English synonyms (e.g. _Volk_ for the 'people') slightly changes the concept; instead of measuring populist attitudes, we introduce nativist as well as populist attitudes. Hence, by careless translations we run the risk of colluding the measure.

[^1]: A scale is measurement invariant across groups if subjects with identical levels of the latent construct have the same expected raw-score on the measure [@drasgow1985equivalence].
[^2]: All analytical material for replication can be found in our online Research Compendium: https://anonymous.4open.science/r/wording-experiment-populist-attitudes-F3B2/README.md (anonymized for the review process).

# A Civic versus Ethnic Conception of "the People"
To test whether an ethnic or civic framing of the people affects respondents' agreement with key populist items we conduct a wording experiment. 
Wording experiments have been used in political science to assess how different question phrasing affects respondents' perceptions of concepts, such as issue ownership [@walgrave2016], climate change [@schuldt2011], and party identity [@sanders2002].
We rely on a nationally representative sample of German citizens over 18 years old (`N=7,034`). 
Given the more pronounced ethnic connotation with the concept of 'people' or _Volk_ in the German language, the German case is particularly suitable to tease out differences between ethnic and civic conceptions of the people in key populist attitude items. 
Our sample is from an online panel provided by the survey company Respondi -- for an overview of all variables in the study, as well as the descriptive information, see Online Appendix (OA pp. A2-A6). 
Quotas for age, gender, and education were included to ensure the representativeness of the sample. 
We have `3,507` respondents in the group that received the ethnic phrasing and `3,527` respondents in the group with the civic phrasing. 
This large sample size allows us to detect also small effects of question wording differences.[^3]

[^3]: A power analysis is conducted and reported on our [Online Research Compendium](https://anonymous.4open.science/r/wording-experiment-populist-attitudes-F3B2/README.md).

```{r dv, topcaption=TRUE}
source(here::here("src/lib/functions.R"))
df <- tibble(Variable = c("POP1","POP2","POP3",
                         "POP4","POP5","POP6"), 
            `Wording ENG` = c("The politicians in the German Bundestag need to follow the will of the [people/citizens].",
               "The [people/citizens], not politicians, should make the most important political decisions",
               "I would rather be represented by an ordinary citizen than by a professional politician.", 
               "The political differences between elites and the people are greater than the differences among citizens",
               "Elected officials talk too much and take too little action",
               "What people call ‘compromise’ in politics is really just selling out on one’s principles"),
            `Wording DE` = c("Die Abgeordneten des Deutschen Bundestags müssen dem Willen [der Bürger/des Volkes] Folge leisten", 
                             "[Die Bürger/Das Volk], und nicht die Politiker, sollte(n) die wichtigsten politischen Entscheidungen treffen",
                             "Ein Bürger würde besser meine Interessen vertreten als ein Berufspolitiker",
                             "Die politischen Unterschiede zwischen den Eliten und [den Bürgern/dem Volk] sind größer als die Unterschiede (zwischen Bürgern)",
                             "Die Politiker reden zu viel und machen zu wenig",
                             "Was in der Politik Kompromiss genannt wird, ist in Wirklichkeit nur ein Verrat von Prinzipien"))

kbl(df, booktabs =T, caption = "\\label{tab:popatt}Akkerman et al Populist Attitudes Scale") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
   column_spec(1, width = "1.5cm") %>%
   column_spec(2, width = "6cm") %>%
   column_spec(3, width = "6cm")

```
```{r results-h1,out.width = "85%", fig.align = 'center', fig.cap = "\\label{fig:results-h1}Differences in Ethnic and Civic Conceptions of the People"}
knitr::include_graphics(here("report/figures", "h1-1.png"))
```
To examine whether civic conceptions of the people elicit different responses than ethnic concepts, we experimentally vary the wording of two of the six items in the scale proposed by @akkerman2014, see Table \ref{tab:popatt}. 
In the original Akkerman et al. scale deployed in the Netherlands, the items `POP1` and `POP2` were phrased in ethnic terms (i,e., with reference to _het volk_). 
We randomize whether respondents received a ethnic framing (i.e., with reference to 'the people; or _das Volk_) or a civic framing (i.e., with reference to 'citizens', or _Bürger_). 
A balance test, reported in our [Online Research Compendium](https://anonymous.4open.science/r/wording-experiment-populist-attitudes-F3B2/README.md), demonstrates that the allocation of the treatment was fully random.
The upper two items in Table \ref{tab:popatt} shows the differences in wording for the two items `POP1` and `POP2` in English and German. 
All the other items were kept constant.[^4].

[^4]:Our intention was to also vary the wording of the item `POP4`. Yet, due to an unfortunate translation error in Qualtrics, we cannot use the item `POP4` in our analysis. However, robustness checks in the Online Appendix (OA) Table A.8 shows that removal of the item does not strongly affect the overall latent construct.

Figure \ref{fig:results-h1} shows the results of a difference of means test for the two treatment groups (ethnic vs. civic wording) for both items -- Table A.5 in the OA shows the coefficients. There is a statistically significant difference in respondent agreement with the ethnic and civic conception of the people for the item `POP2` (''the people, not politicians, should make the most important political decisions''), but not for `POP1` (''The politicians need to follow the will of the people'').
The negative coefficient for `POP2` indicates that when 'the people' is conceived as _das Volk_ respondents are more likely to agree with the item's statement compared to the civic wording of _Bürger_.
Yet, as the results show, the difference is small in substantive terms.

[^5]: Our [Online Research Compendium](https://anonymous.4open.science/r/wording-experiment-populist-attitudes-F3B2/README.md) demonstrates that there are no other socio-demographic variables that mediate the effect.

_How does the effect of an ethnic conception of the people vis-à-vis a civic conception vary across respondents?_
The left-hand panel of Figure \ref{fig:results-h2} shows the results of an OLS regression analysis[^6] with the an interaction between the treatment and a variable measuring whether or not the respondent voted for the far right party _Alternative für Deutschland_ -- for more details see Table A.6 in the OA).
We see that far right voters show much more agreement to items `POP1` and `POP2` when it is phrased in an ethnic way. 
These effects are substantial and statistically significant. 
For the item `POP2` (''the people, not politicians, should make the most important political decisions''), when voting for the AfD and being in the ethnic conception of the people condition, one scores on average a full point higher on a 5-point scale, which is about a full standard deviation.
For the item `POP1` (''The politicians need to follow the will of the people''), the effect size is about half a point (i.e. half a standard deviation).

[^6]: As specified in the Pre-Analysis Plan, we only include unbalanced covariates. OA p. A-7 demonstrates the balance checks, demonstrating that only political interest and income are unbalanced, and therefore included as a covariate.

```{r results-h2, out.width = "100%", fig.align = 'center', fig.cap = "\\label{fig:results-h2}Average Marginal Effects for Ethnic Conception of the People"}
knitr::include_graphics(here("report/figures", "h2-1.png"))
```

In a next step, we examine whether respondents with lower and higher levels of exclusive national identity are affected differently by the treatment. 
Inspired by @mader2021, we measure exclusive national identity using a four-item additive scale on a 5-point Likert scale, which asks respondents to rate how important the following aspects are to be considered German: being born in Germany (a), having German ancestors (b), being able to speak German (c), and sharing share German manners and norms (d). 
The right-hand panel of Figure \ref{fig:results-h2} shows the interaction between the treatment and the exclusive national identity variable on the two items -- for more details see Table A.7 in the OA). 
Compared to not adhering to an exclusive national identity, the average score of fully adhering to an exclusive national identity is 0.2 higher on a 5-point scale for the item `POP2` (''the people, not politicians, should make the most important political decisions'').
Compared the standard deviation of this item (1.17, see Table A.4 in the OA), this is a small effect.
For the other item, the line is much less steep over the various levels of adherence to an exclusive national identity.
Overall, we see that the difference between an ethnic and civic wording is only statistically significant for respondents with higher levels of exclusive national identity (ca. `>2.5` on the exclusive national identity scale). 
Both the analyses from the left and right-hand panel of \ref{fig:results-h2} suggest that far right voters and respondents with an exclusive national identity interpret 'the people' (_das Volk_) in an ethno-cultural way. 
The results of Figures \ref{fig:results-h1} and \ref{fig:results-h2} seem to suggest that the assumption of measurement invariance is not met, as we find statistically significant difference for the `POP2` (''the people, not politicians, should make the most important political decisions'') item.

In addition, we have explored how the effect of an ethnic conception of the people vis-à-vis a civic conception vary across age and region. People that are born during or closely after WWII  as well as people born and/or living in certain regions in Germany (e.g. the East or Bavaria) could potentially have a different reaction to the two conceptions of the people -- all figures can be found in the Online Appendix (Figure A-2, p. A9) as well as in our Online Research Compendium.
**_Age_.** We demonstrate in the upper-left panel of Figure A-2 in the OA that for younger age groups (i.e. all age groups younger than 50 years old) there is no difference for the civic and ethnic conception of the people. Yet, for the older population (50 years old and more) we see that people receiving the ethnic treatment report higher levels of agreement with the items compared to people receiving the civic treatment for both `POP1` (''The politicians need to follow the will of the people''), the effect size is about half a point (i.e. half a standard deviation) and `POP2` (''the people, not politicians, should make the most important political decisions''). The effect size is still small in substantive terms: between a `0.1` and `0.3` increase on a 5-point scale.
**_Region_.** We show in the upper-right panel of Figure A-2 in the OA that while for most regions in Germany, there is no statistically significant difference between receiving an ethnic or a civic treatment. Yet, for the regions Sleswig-Holstein (i.e. the region bordering Denmark) and Hessen (mid-Germany), we see that people receiving the ethnic treatment report lower levels of agreement with the items compared to people receiving the civic treatment for both `POP1` and `POP2`. The same effect holds for the region Bremen for `POP1` and for Berlin and North-Rhine Westphalia for `POP2`. This indicates that in Western districts, people are sensitive to the ethnic conception of the 'people' and all else equal, under-report their populist attitudes by approximately `0.5` points on a 5-point scale.

We also explored whether the difference in civic and ethnic conception is only for voters of far right parties or also exists for people placing themselves on the edges of the ideological spectrum. An interaction between the treatment and ideological self-placement -- for more details see lower-left panel of Figure A-2 in the OA) -- demonstrates that for `POP1` (''The politicians need to follow the will of the people'') there is a positive, yet insignificant, relationship between ideology and receiving the ethnic treatment; i.e. the more right-wing, the higher your score on the item when receiving the ethnic treatment compared to the civic treatment. For `POP2` (''the people, not politicians, should make the most important political decisions''), we see the opposite: Left wing people report higher scores on this item when shown the ethnic conception of the people compared to the civic conception. Yet, the effect is substantially small, between approximately a `0.2` and `0.05` decrease on a 5-point scale. This effect is negative, meaning that the more right-wing, the smaller the difference between the score for the item for the two treatments. The effect becomes insignificant for an ideology score of 7 and higher. In addition, we explored the differences between those who are opponents and proponents of political pluralism, as the civic conception of the people might invite a more pluralistic view on who are included in the people. The lower-right panel of Figure A-2 in the OA demonstrates the opposite: Respondents who are negative towards political pluralism (i.e. low scores) that receive the ethnic treatment report lower levels of agreement with the items compared those who are positive towards political pluralism (i.e. high scores)  for both `POP1` and `POP2`. The effect is bigger for `POP2` than for `POP1`.

_Does an ethnic and civic framing of  the items `POP1` and `POP2` affect the measurement invariance assumption of the latent construct of populist attitudes?_
To answer this question, we cannot just compare the means of the group, since we would assume that the measures function similar in these different groups [e.g. see @hirschfeld2014improving; @van2021methods; @bieda2017universal]. We therefore need to test the degree to which measures are invariant across groups. To do so, Multiple-Group Confirmatory Factor Analysis (MG-CFA) has become the de-facto standard in psychology [@chen2008happens].
We first test the measurement variance in a Confirmatory Factor Analysis (see Table A.9 in the OA). 
The averages of the civic and ethnic conception scales of populist attitudes differ significantly:
Respondents exposed to the civic conception score higher (`0.44` on a 5-point scale) on the populist attitudes scales than respondents in the ethnic conception condition. 
In terms of model fit, both constructs show very similar goodness-of-fit statistics (i.e., CFI, BIC, RMSEA, and SRMR). Interestingly, however, both the 'ethnic' and the 'civic' scale show a satisfactory model fit (`CFI > 0.9`; `RMSEA < 0.08`; and `SRMR < 0.02`) -- for more information see OA Table A.8. 
Moreover, when testing metric and scalar invariance -- i.e. is the magnitude of loadings similar across groups or are the magnitude of loadings and the intercept difference across groups -- we find statistical significant Chi-square differences for the metric invariance and scalar invariance models, shown in Table \ref{tab:mi-cfa}.
This indicates that the assumption that the different conceptions of the people function similarly is not met.

```{r mi-cfa, topcaption=TRUE}
load(here("data/intermediate/cleaned_experiment.RData"))
source(here("src/analysis/data-for-analyses.R"))
source(here("src/analysis/explorative_cfa.R"))

mi %>%
  select(`Equivalence Tests` = Models, AIC, BIC, `Chisq diff`, `Pr(>Chisq)`) %>%
   filter(`Equivalence Tests` != "Configural Invariance") %>% 
  kbl(booktabs =T, caption = "\\label{tab:mi-cfa}Measurement Invariance - CFA Models") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = "4cm") %>%
  column_spec(2, width = "2cm") %>%
  column_spec(3, width = "2cm") %>%
  column_spec(4, width = "2cm") %>%
  column_spec(5, width = "2cm")   
```

Additionally, we have estimated a graded Item Response Theory (IRT) analysis [see also @vanhauwaert2020].
This scaling model shows that `POP1` and `POP2` contribute slightly more information to the latent construct when the items are ethnically phrased, as opposed to a civic framing (see Figure A.5 and Table A.13 for more information in the OA). 
While the means of the scales with the different conceptions of the people are not different, testing metric and scalar invariance -- see Table \ref{tab:mi-irt} -- we see that the equal slope assumption (metric invariance) as well as one form of scalar invariance (i.e. the one with free means) are not met.
This indicates that even varying the wording of 'the people' in two items of the full @akkerman2014 item battery is enough to demonstrate that the scales are not equivalent anymore.

```{r "irt-code", fig.align="center", fig.height=8, fig.width=12, echo=FALSE, message=FALSE, warning=FALSE,results = 'hide'}
# Measurement Invariance
library(mirt)
dat <- matrix(data = NA, nrow = dim(d)[1], ncol = 5)
dat[,1] <- d$POST_1
dat[,2] <- d$POST_2
dat[,3] <- d$POST_3
dat[,4] <- d$POST_5
dat[,5] <- d$POST_6
colnames(dat) <- c("item1", "item2", "item3", "item4", "item5")
group <- d$ethnic
group <- if_else(group==0, "C1", "E1")

mod_configural <- multipleGroup(dat, 1, group = group)
mod_metric <- multipleGroup(dat, 1, group = group, invariance=c('slopes')) #equal slopes
mod_scalar2 <- multipleGroup(dat, 1, group = group, #equal intercepts, free variance and means
                             invariance=c('slopes', 'intercepts', 'free_var','free_means'))

a1 <- anova(mod_metric, mod_configural) #equal slopes only
a2 <- anova(mod_scalar2, mod_metric) #equal intercepts, free variance and mean

dat <- a1[2,] %>% 
  add_case(a2[2,]) %>% 
  mutate(`Equivalence Tests` = c("Metric Invariance",
             "Scalar Invariance")) %>% 
  dplyr::select(`Equivalence Tests`, AIC, BIC, `LogLikelihood` = logLik, `Significant Difference` = p)
```
```{r "irr-code2"}
dat %>%
  kbl(booktabs =T, caption = "\\label{tab:mi-irt}Measurement Invariance - IRT Models") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = "4cm") %>%
  column_spec(2, width = "2cm") %>%
  column_spec(3, width = "2cm") %>%
  column_spec(4, width = "2cm") %>%
  column_spec(5, width = "4cm") 
```

Moreover, we have created an additive scale (see Table A.11 in the OA), as well as constructed a scale using the approach of @wuttke2020whole  (see Table A.12 in the OA) -- i.e. using the lowest value on one of the items that construe the latent construct.
For both scales, we find statistical significant average values for the scales based on ethnic and civic conception of the people.
For both scaling methods, people in the civic conception have on average a higher mean value compared to the ethnic conception.
While we are not able to formally test all forms of measurement invariance, a statistically significant mean indicates that the measurements are not equivalent.

# Conclusion
The populist attitude construct aims to operationalize populism as a thin-centered ideology on the individual level [@akkerman2014; @hawkins2012]. From this perspective, populist ideation is present both on the left and the right side of the political spectrum. The concept 'the people' is an empty signifier that is discursively filled in different ways by left- and right-wing populists. Therefore it is important that the 'populist attitudes' construct is not biased toward one group or the other. Yet, existing populist attitudes scales have been inconsistent in using 'ethnic' or 'civic' conceptions of the people. Moreover, in many languages, the concept of `the people` has a clear ethnic connotation. This could hinder individual-level measurement of populism as a thin-centred ideology.

Using a wording experiment, we therefore examined whether German respondents presented with an ethnic framing of 'the people' (_das Volk_) answered key populist attitude items differently than respondents who recevied an ethnic framing of `the people' as citizens (_Bürger_).

We find that the measurement invariance assumption does not hold: 
Using different wordings does not lead to equivalent measaures.
Moreover, there are statistically significant differences between the two wordings. 
We only varied the ethnic and civic conception for two items.
It is likely to be more consequential in survey batteries that predominantly include items contrast 'the people' with elites to measure populist attitudes [for an overview of different batteries, see @castanho2020]. 
The populist attitude scales proposed by @schulz2018 and @stanley2011, for instance, contain references to 'the people' in 7/9 and 7/8 of the items, respectively.

While respondents receiving the items with an ethnic conception of the people scored overall higher on the items, their overall score on the populist attitude scale was not higher (shown in OA pp.A10--A13). The higher levels of agreement with the items displaying an ethnic conception of the people is particularly pronounced for voters of the far right (shown in the left panel of Figure 2), as well as for older voters (shown in OA pp.A8--A9). The difference between an ethnic and a civic conception of the people in the items was also only significant for respondents who adhered a moderate to high level of exclusive national identity (shown in OA pp.A8--A9). This finding is highly suggestive of the 'ethnic' interpretation of `the people' -- at least in the German language -- thereby conflating populism with nationalism.

Importantly, the degree to which 'the people' is understood in denoting a specific ethno-cultural group likely differs across countries.
This type of differential item interpretation might therefore lead to even more measurement invariance when used in cross-national research [@castanho2020]. 
In other words, if the ethno-cultural connotation of the the people varies across countries, this may hamper cross-national application of populist attitude scales that include multiple items referring to 'the people'. 
In fact, in a comparison of different populist attitudes scales @castanho2020 find that the scales by @schulz2018 and @stanley2011 perform worse than the scale proposed by @akkerman2014. 
To empirically establish whether differential interpretations of `the people' matters for measurement invariance, it would be worthwhile replicating this study for different countries. In addition, it would be worthwhile in replicating this analysis for different scales than the @akkerman2014 scale.

Crucially, our findings suggest that populist attitude scales that rely on ethnic measures of 'the people' might not be entirely appropriate for measuring populism as thin-ideology that attaches itself to both inclusive and exclusionary ideologies. What is more, While (ethnic) nationalism and populism often go hand in hand, it is important that we keep both concepts analytically distinct, both conceptually and empirically [@bonikowski2017ethno]. An ethnic priming of 'the people' in the populist attitudes scale, makes it more difficult to keep ethnic nationalism and populism distinct.

One could argue theoretically that 'the people' captures the meaning of the concept of populism better than 'citizens'. 
After all, it populism assumes the people to be a homogeneous entity. 
Yet, our findings suggest that also a civic framing of the concept yields a satisfactory model fit when estimating a Confirmatory Factor Analysis. 
In addition, our results suggest that an ethnic framing of the citizens runs the risk overestimating the degree of populist attitudes among more nativist citizens. 
To avoid priming respondents with an ethnic conception of the people, future studies should use a consistent ethnic or civic conception of the people when measuring populist attitudes, bearing in mind that the civic conception does lead to higher values on the scale.

What is more, our results suggest that political scientists should be careful when commissioning translations of survey items. While different translations of key concepts may seem equivalent to non-specialists, our findings show that such nuances can yield measurement invariance - impeding comparison across countries.

\newpage
# References